Article Title

### 2013-12-06 plos.tex
Updated Abstract and Author summary, Figure 7.

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.  It is anticipated that correlations measured in such recordings will uncover important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: the biasing of the estimate towards a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies.  Therefore, the selection of the most efficient estimator is an empirical question that depends on the system under investigation.

In this study, we sought to identify the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals we recorded the activity of nearly every neuron in volumes of cortex about 100 $\mu$m in diameter (150--350 cells).  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with several shared latent factors to capture unobserved inputs and network fluctuations.  To test this hypothesis, we demonstrated that, in cross-validation tests, a covariance matrix estimator based on this correlation structure outperformed estimators regularized toward several other plausible low-dimensional approximations. We analyzed how the inferred network of interactions was related to the physical distances between cells and their orientation preferences.  Consistent with known synaptic connectivity, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and local field potentials.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlation matrices of the activity of populations of neurons are useful descriptors of their functional organization with implications for stimulus coding and circuit architecture.  Estimates of the correlation matrix can be dramatically improved by \emph{regularization}: the biasing of the estimate toward one of several possible low-dimensional approximations.  Using synthetic data, we demonstrated that different estimators performed better or worse than others depending on how well their regularization targets captured the structure of the true correlation matrix.  Then, we compared the performance of several estimators on the population calcium activity of relatively large and dense groups of neurons in mouse visual cortex under visual stimulation.  We found that the best estimator for such neural data was composed of a sparse network of partial correlations between pairs of neurons combined with several latent units interacting with the entire population. We hypothesize that such succinct representations of the correlation structure may more accurately reflect the circuit's functional organization than the conventional correlations.  
\section*{Introduction}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect aspects of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

A correlation matrix provides more information than the equivalent number of pairwise correlations assessed in isolation.  Yet early studies of neural correlations were based on measurements from isolated pairs of neurons; the effects of correlations on neural coding were then extrapolated in simulations and theoretical studies \cite{Shadlen:1998,Zohary:1994}.  Modern multineuronal recordings from increasingly large and dense subsets of neurons in intact circuits allow estimating the entire correlation matrix of large populations of highly interconnected neurons.  Such estimates can be transformed into various representations that express different aspects of the correlation structure and may suggest different mechanistic interpretations.  For example, factor analysis or principle component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.  The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.  Under several assumptions (predominance of linear effects ( sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons. In multivariate normal distributions, zero partial correlations between pairs of variables indicate conditional independence or lack of interaction between the pair and the network of non-zero interactions constitute Gaussian Graphical Models or Gauss-Markov Random Fields \cite{Koller:2009}. With other distributions, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}. Estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to map gene interactions \cite{Schafer:2005,Peng:2009} or to establish functional connectivity between brain regions from fMRI data \cite{Varoquaux:2012,Ryali:2012}.

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations measured in such recordings are anticipated to uncover important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: biasing of the estimate towards a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies present in the system.

In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations. Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.  Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure closely matches the real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that proved most efficient comprised a sparse network of partial correlations between pairs of neurons combined with several common latent factors, with both components required for efficient estimation. Because of its superior benefit for estimation, we proposed that this inferred structure could prove more relevant for the description of functional connectivity than the usual correlation matrix in densely sampled neural recordings. As a first application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.

\section*{Introduction}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

By their construction, the correlation between a pair of cells expresses the degree to which the activity of one cell predicts the activity of the other cell. With multiple cells If interactions between variables have a strong linear components, then correlations also statistical dependence.

A correlation matrix provides more information than the equivalent number of pairwise correlations assessed in isolation.  Yet early studies of neural correlations were based on measurements from isolated pairs of neurons; the effects of correlations on neural coding were then extrapolated in simulations and theoretical studies \cite{Shadlen:1998,Zohary:1994}.  Modern multineuronal recordings from increasingly large and dense subsets of neurons in intact circuits allow estimating the entire correlation matrix of large populations of highly interconnected neurons.  Such estimates can be transformed into various representations that express different aspects of the correlation structure and may suggest different mechanistic interpretations.  For example, factor analysis or principal component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.  The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.  Under several assumptions (predominance of linear effects ( sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons. In multivariate normal distributions, zero partial correlations between pairs of variables indicate conditional independence or lack of interaction between the pair and the network of non-zero interactions constitute Gaussian Graphical Models or Gauss-Markov Random Fields \cite{Koller:2009}. With other distributions, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}. Estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to map gene interactions \cite{Schafer:2005,Peng:2009} or to establish functional connectivity between brain regions from fMRI data \cite{Varoquaux:2012,Ryali:2012}.
